---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r ,results='hide'}
data<-read.csv("activity.csv",sep=",",header=T)
head(data)
summary(data)
str(data)
```


## What is mean total number of steps taken per day?
```{r ,results='hide'}
library("dplyr")
library("ggplot2")
```

####Total number of steps taken per day
```{r}
by.day<-group_by(data, date)
steps.day<-summarise(by.day, total=sum(steps,na.rm=T))
steps.day
```

####Calculate the median and mean of the total number of steps taken per day.
```{r}
summary(steps.day)
```
Mean is 9.354 steps.  
Median is 10.395 steps

####Histogram of the total number of steps taken each day
```{r histogram}
qplot(total, data=steps.day, main="Total number of steps taken each day",
    xlab="Number of steps")
```


## What is the average daily activity pattern?

####A time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r plot}
by.interval<-group_by(data, interval)
steps.series<-summarise(by.interval, meanstep=mean(steps, na.rm=T))
qplot(interval,meanstep,data=steps.series,geom="line",main="Average daily 
      activity pattern", xlab="5 minutes interval",ylab="Average number of steps
      (all days)")
```

####5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps.
```{r}
filter(steps.series, meanstep==max(meanstep))
```


## Imputing missing values

####Total number of missing values in the dataset
```{r}
summary(data)
sum(is.na(data$steps))
```

####Strategy for filling in all of the missing values in the dataset
Use the mean for that 5-minute interval.

####New dataset that is equal to the original dataset but with the missing 
data filled in.
```{r}
datnew<-data
datnew$mean<-steps.series$meanstep
lost<-is.na(datnew$steps)
datnew[lost,"steps"]<-datnew[lost,"mean"]
head(datnew)
```

####Histogram of the total number of steps taken each day.
```{r}
by.day2<-group_by(datnew,date)
step.day2<-summarise(by.day2, sum=sum(steps))
qplot(sum, data=step.day2,main="Total number of steps taken each day (II)",
      xlab="Number of steps")
```

####Mean and median total number of steps taken per day.
```{r}
summary(step.day2)
```

####Impact of imputing missing data
Decreases the number of days with few steps.  
Increases the median and mean.  
Significantly increase days with about 11.000 steps (around median and mean)


## Are there differences in activity patterns between weekdays and weekends?

####Create a new factor variable in the dataset with two levels – “weekday” and 
“weekend”
```{r}
datnew$dayweek<-weekdays(as.Date(datnew$date))
datnew$dayweek[datnew$dayweek=="sábado"|datnew$dayweek=="domingo"]<-"weekend"
datnew$dayweek[!(datnew$dayweek=="weekend")]<-"weekday"
unique(datnew$dayweek)
```

####Panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r plot2}
by.interval1<-group_by(datnew, dayweek, interval)
step.series1<-summarise(by.interval1, meanstep1=mean(steps))
qplot(interval,meanstep1, data=step.series1,facets=dayweek~. ,geom="line", 
      main="Activity patterns: weekdays and weekends", xlab="5 minutes interval",
      ylab="Average number of steps").
```
